--- 
title: |
  | My Dissertation 
  |
date: | 
  | Malavika Rajeev
  |
  |
  |
  | A DISSERTATION PRESENTED TO THE FACULTY 
  | OF PRINCETON UNIVERSITY 
  | IN CANDIDACY FOR THE DEGREE OF DOCTOR OF PHILOSOPHY
  |
  |
  |
  | [REPLACE THIS WITH ACCEPTANCE STATEMENT AFTER COMMITTEE MEETING] 
  |
  |
  |
  | Advisor: C. Jessica E. Metcalf
  |
  |
  |
  | `r format(Sys.Date(), '%B %Y')`
documentclass: book
lof: True
lot: True
biblio-style: apalike
bibliography_main: [bib/book.bib]
bibliography_supp: [bib/packages.bib]
link-citations: yes
description: "WHere does this show up exactly?"
site: bookdown::bookdown_site
---

\mainmatter

\pagestyle{fancy}
\lhead{}
\chead{}
\rhead{\thepage}
\cfoot{} 
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

# Introduction {-} 

This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

The **bookdown** package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.org/tinytex/>.

```{r include=FALSE}
# # automatically create a bib database for R packages
# knitr::write_bib(c(
#   .packages(), 'bookdown', 'knitr', 'rmarkdown'
# ), 'packages.bib')
```
